There have been a number of developments recently moving towards
planning under uncertainty using systems that were intended for
sequential planning in deterministic problems.  Notably for example,
\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
probabilistic track of the 2004 International Planning Competition.
In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
plans and corresponding execution traces.
%%
More computationally expensive approaches in this vein combine
sampling strategies on valuations over {\em runtime variables} with
deterministic planning procedures. The outcome is typically a more
robust sequential plan~\cite{yoon:etal:2008}, or contingent
plan~\cite{majercik:2006}. 

%%No! They simply haven't been evaluated in PO settings. They may, or
%%may not struggle. They have sampling of traces, and that would
%%include observations, and therefore evolutions of beliefs. SSAT was
%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
%%suited to POMDPs.

% Normally if we are going to compare with related work, we do
% actually *compare*. Why didn't you try those approaches? I think
% it's safe to say that FFR will struggle. Why would it even include
% in its plan an observational action that doesn't change the world?

%%
%% However, as we said in the introduction,
%% all these approaches struggle in partially observable domains as they
%% rely on being able to determine the state at all times.

Also leveraging deterministic planners in problems that feature
uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
---i.e., sequential planning in unobservable worlds--- can be modelled
as a deterministic problem, and therefore solved using sequential
systems. In this conformant setting, advances have been towards
compact representations of beliefs amenable to existing best-first
search planning procedures, and lazy evaluations of beliefs. Most
recently this research thread has been extended to contingent planning
in fully observable non-deterministic
environments~\cite{albore:etal:2009}.
%%
The continual planning system that motivated our
project~\cite{brenner:nebel:jaamas09} also has this characteristic,
and has been applied in completely observable domains, particularly
those featuring multiple communicating agents. The use of knowledge
operators in domains allows plans that act to gain knowledge, but the
approach assumes that such actions are deterministic and reliable, an
assumption that we relax in this work.

Our work is motivated by domains that contain a mixture of task
planning and observation planning. There have been a number of recent
papers representing observation planning problems as POMDPs and using
various techniques to manage the large state
space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
approach in a vision algorithm selection problem. In their case there
is a natural hierarchical decomposition of the problem which allows
them to solve large problems by breaking them into a set of small
POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
represent a preference elicitation problem as a POMDP and take
advantage of symmetry in the belief space ---essentially the idea that
it does not matter what the value of the variable you are trying to
observe is--- to exponentially shrink the state space. Although we
have been actively exploring the \citeauthor{doshi08:pref_elic}
approach, we have yet to find those exploitable structures in our
domain due to the task planning requirement.
